---
title: "EDA dashboard"
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'data_dashboard.html'))})
output: 
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 1
---

```{r libraries, message=FALSE, warning=FALSE, include=FALSE}
library(readxl)  # to import the data in xlsx file
library(finalfit)  # for function ffglimpse(), extension of summary
library(ggplot2)  # for plotting
library(plyr)  # for counting the frequency from histogram
library(ggpubr)  # to get the function ggarrange; to combine ggplots
library(cowplot)  # to draw density histogram at the sides of correlation plot

# The dataset
pop_raw = readxl::read_xlsx("Tested_Poplar.xlsx",
                            na = "NA")  # defining the code for missing data used in the file
# Getting only the important variables
wild_yr2000 = pop_raw[,c("Classe",  # the type of population source 
                         "Population",  # population of origin
                         "Clone",  # the observed poplar individuals
                         "Notemax.2010 Noveltree",  # field resistance 
                         "latence.93JE3.2010",   # latent period
                         "nbsores.93JE3.2010",  # number of spores
                         "taille.93JE3.2010")]  # size of spores

colnames(wild_yr2000)[4] = "Notemax.2010"  # because the name was too long

# Check the class of each variable
str(wild_yr2000)

# Coercing
## Classe = chr to factor
## Population = chr to factor
## Clone = chr to factor
cols = c("Classe", "Population", "Clone")  # cols to be coerced

wild_yr2000[cols] = lapply(wild_yr2000[cols], factor)  # coerced to factor

# New data structure
str(wild_yr2000)

rownames_remove <- c("témoin", "coll. Nationale")

wild_yr2000_clean = subset(wild_yr2000,!(Classe %in% rownames_remove))
```

# Obj 1 Data dashboard for resistant vs susceptible (in the field)
Turned out the NA is due to  
1. Non-measured field resistance in the poplars from national collection.   
2. Non-active infection in certain controls.  

```{r res vs sus field}
# Resistant poplars in the field: Assigning -5 to NA to avoid confusion in R
wild_yr2000_clean[is.na(wild_yr2000_clean)] = -5

# Add a new variable: resistant vs susceptible
# Loop, -5 = resistant, != -5 susceptible
wild_yr2000_clean[,8] = "NA"

names(wild_yr2000_clean)[8] = "Res_vs_sus_note"

for (i in c(1:nrow(wild_yr2000_clean))) {
  if(wild_yr2000_clean$Notemax.2010[i] == -5)
    wild_yr2000_clean[i,8] = "Resistant"
  
  if(wild_yr2000_clean$Notemax.2010[i] != -5)
    wild_yr2000_clean[i,8] = "Susceptible"
}

# Plotting resistant vs susceptible poplars in terms of field resistance
(resist_field = ggplot(wild_yr2000_clean, aes(x = Res_vs_sus_note)) +
  geom_bar(aes(fill = Res_vs_sus_note)) +
  geom_text(stat = "count", aes(label = ..count..), vjust = -0.5) +
  labs(title = "Resistant vs susceptible poplars",
       x = "Resistant vs Susceptibe",
       y = "Counts") +
  scale_fill_manual(values = c("#87CEEB", "#FF6A6A"),
                    breaks = c("Resistant", "Susceptible"),
                    name = "Resistant vs Susceptible",
                    labels = c("Resistant", "Susceptible")) +
  theme_classic() +
  theme(axis.title.x = element_blank()))

# Checking the population of origin of the resistant poplars
pop = plyr::count(wild_yr2000_clean$Population[
  which(wild_yr2000_clean$Notemax.2010 == -5)
])

names(pop)[1] = "Population"

clone = plyr::count(wild_yr2000_clean$Clone[
  which(wild_yr2000_clean$Notemax.2010 == -5)
])

names(clone)[1] = "Clone"

pop$Clone = paste(clone$Clone)

pop

```

It seems that there is only one poplar (71072-501) from St Genix that is resistant 
in the field.

Let's see the resistant poplars in terms of lab resistance
Those with latence = 14 and taille = -5

```{r res vs sus lab}
# Add a new variable: resistant vs susceptible (lab)
# Loop, -5 = resistant, != -5 susceptible
wild_yr2000_clean[,9] = "NA"

names(wild_yr2000_clean)[9] = "Res_vs_sus_lab"

for (i in c(1:nrow(wild_yr2000_clean))) {
  if(wild_yr2000_clean$taille.93JE3.2010[i] == -5
     & wild_yr2000_clean$latence.93JE3.2010[i] == 14)
    wild_yr2000_clean[i,9] = "Resistant"
  
  if(wild_yr2000_clean$taille.93JE3.2010[i] != -5
     & wild_yr2000_clean$latence.93JE3.2010[i] != 14)
    wild_yr2000_clean[i,9] = "Susceptible"
}

# Plotting resistant vs susceptible poplars in terms of lab resistance
(resist_lab = ggplot(wild_yr2000_clean, aes(x = Res_vs_sus_lab)) +
  geom_bar(aes(fill = Res_vs_sus_lab)) +
  geom_text(stat = "count", aes(label = ..count..), vjust = -0.5) +
  labs(title = "Resistant vs susceptible poplars",
       x = "Resistant vs Susceptibe",
       y = "Counts") +
  scale_fill_manual(values = c("#458B74", "#EE3B3B"),
                    breaks = c("Resistant", "Susceptible"),
                    name = "Resistant vs Susceptible",
                    labels = c("Resistant", "Susceptible")) +
  theme_classic() +
  theme(axis.title.x = element_blank()))

# Checking the population of origin of the resistant poplars (lab)
pop2 = plyr::count(wild_yr2000_clean$Population[
  which(wild_yr2000_clean$Res_vs_sus_lab == "Resistant")
])

names(pop2)[1] = "Population"

clone2 = plyr::count(wild_yr2000_clean$Clone[
  which(wild_yr2000_clean$Res_vs_sus_lab == "Resistant")
])

names(clone2)[1] = "Clone"

pop2$Clone = paste(clone2$Clone)

pop2
```

Seems that there are different poplars that are resistant in field and in the lab.

---

# Obj 2 Average resistance. Data dashboard for summary statistics
```{r summary statistics, fig.height=5, fig.width=10, message=FALSE, warning=FALSE}
# Field resistance (notemax)
# Removing -5 because they're outliers
wild_yr2000_new = wild_yr2000_clean[which(wild_yr2000_clean$Notemax.2010 != -5 &
                                      wild_yr2000_clean$taille.93JE3.2010 != -5), ]

summary(wild_yr2000_new)

NM = ggplot(wild_yr2000_new, aes(x = Notemax.2010)) +
  geom_histogram(binwidth = 0.5) +
  geom_density(aes(y = 0.3 * ..count..)) +
  geom_vline(aes(xintercept = mean(Notemax.2010), color = "Mean"), size = 1) +
  geom_vline(aes(xintercept = median(Notemax.2010), color = "Median"), size = 1) +
  geom_vline(aes(xintercept = quantile(Notemax.2010, prob = 0.25), color = "Q1"), 
             size = 1) +
  geom_vline(aes(xintercept = quantile(Notemax.2010, prob = 0.75), color = "Q3"), 
             size = 1) +
  labs(title = "Field resistance",
       caption = "number of samples = 72, sd = 0.63",
       x = "Score",
       y = "Counts") +
  scale_color_manual(name = "Statistics", values = c(Mean = "red", Median = "orange", Q1 = "blue",
                                                     Q3 = "blue")) +
  xlim(1, 6) +
  ylim(0, 40)
 
# Latent period (latence)
LP = ggplot(wild_yr2000_new, aes(x = latence.93JE3.2010)) +
  geom_histogram(binwidth = 0.5) +
  geom_density(aes(y = 0.3 * ..count..)) +
  geom_vline(aes(xintercept = mean(latence.93JE3.2010), color = "Mean"), size = 1) +
  geom_vline(aes(xintercept = median(latence.93JE3.2010), color = "Median"), size = 1) +
  geom_vline(aes(xintercept = quantile(latence.93JE3.2010, prob = 0.25), color = "Q1"), 
             size = 1) +
  geom_vline(aes(xintercept = quantile(latence.93JE3.2010, prob = 0.75), color = "Q3"), 
             size = 1) +
  labs(title = "Latent period",
       caption = "sd = 0.69",
       x = "Days",
       y = "Counts") +
  scale_color_manual(name = "Statistics", values = c(Mean = "red", Median = "orange", Q1 = "blue",
                                                     Q3 = "blue")) +
  xlim(7, 12) +
  ylim(0, 40)

# sd(wild_yr2000_new$latence.93JE3.2010)

# Number of spores
NS = ggplot(wild_yr2000_new, aes(x = nbsores.93JE3.2010)) +
  geom_histogram(binwidth = 1.0) +
  geom_density(aes(y = 0.3 * ..count..)) +
  geom_vline(aes(xintercept = mean(nbsores.93JE3.2010), color = "Mean"), size = 1) +
  geom_vline(aes(xintercept = median(nbsores.93JE3.2010), color = "Median"), size = 1) +
  geom_vline(aes(xintercept = quantile(nbsores.93JE3.2010, prob = 0.25), color = "Q1"), 
             size = 1) +
  geom_vline(aes(xintercept = quantile(nbsores.93JE3.2010, prob = 0.75), color = "Q3"), 
             size = 1) +
  labs(title = "Number of spores",
       caption = "sd = 9.49",
       x = "Number (counts)",
       y = "Counts") +
  scale_color_manual(name = "Statistics", values = c(Mean = "red", Median = "orange", Q1 = "blue",
                                                     Q3 = "blue")) +
  xlim(0, 60) +
  ylim(0, 40)

# sd(wild_yr2000_new$nbsores.93JE3.2010)

# Spore size
SS = ggplot(wild_yr2000_new, aes(x = taille.93JE3.2010)) +
  geom_histogram(binwidth = 0.5) +
  geom_density(aes(y = 0.3 * ..count..)) +
  geom_vline(aes(xintercept = mean(taille.93JE3.2010), color = "Mean"), size = 1) +
  geom_vline(aes(xintercept = median(taille.93JE3.2010), color = "Median"), size = 1) +
  geom_vline(aes(xintercept = quantile(taille.93JE3.2010, prob = 0.25), color = "Q1"), 
             size = 1) +
  geom_vline(aes(xintercept = quantile(taille.93JE3.2010, prob = 0.75), color = "Q3"), 
             size = 1) +
  labs(title = "Spore size",
       caption = "sd = 0.68",
       x = "Score",
       y = "Counts") +
  scale_color_manual(name = "Statistics", values = c(Mean = "red", Median = "orange", Q1 = "blue",
                                                     Q3 = "blue")) +
  xlim(1, 5) +
  ylim(0, 40)

# sd(wild_yr2000_new$taille.93JE3.2010)

# Combining the plots
dist = ggpubr::ggarrange(NM, LP, NS, SS)

dist
```

* Field resistance  
Seems that the poplars are in intermediate resistance, on average. For a note,
field resistance is measured using scoring system from 1 - 6; score 1 means
that infection covers only a small area of leaves, score 6 means that infection
covers almost all area of poplar leaves. In other words, when field resistance
score is 1 then the poplar is resistant, when field resistance score is 6 then
the poplar is susceptible.  
  
  In the field, the score is a little bit more than 3, which means that the infection
moderately covers poplar's leaf area. This being said that the poplars are either
resistant or susceptible in the field.

* Lab resistance, parameter: latent period  
On average, poplars started showing the first infection (or symptoms) on day
8.5 after inoculation. Day 8.5 means that the infection starts on day 8 after 
midday. Seems that the poplars can inhibit the colonization of the pathogen 
until day 8, on average.

* Lab resistance, parameter: number of spores  
Poplar individuals show a variable number of spores. Because of this, there are 
not many individuals that have similar number of spores (see the Y-axis scale). 
It seems that there are 2 peaks in the distribution, showing the segregation
between indivdiuals. This is the indication that the population is made of 2 groups
of populations. On average, the poplars are colonized by 30 spores (max = 56 spores)
even though there are not so many poplars colonized by 30 spores. There are more
poplars colonized by either 25 or 36 spores, indicating that the poplars may be
moderately resistant.

* Lab resistance, parameter: size of spores  
On average, the size of spores is scored closer to 4 than to 3. It shows that the
poplars cannot inhibit the development of the pathogen's spore once the spores
colonize the poplars. It shows that the poplars are rather susceptible for this 
type of pathogen (code: rust 93JE3)

* Overall  
The poplars are closer to being susceptible than to being resistant in the lab 
for this particular type of pathogen (code: rust 93JE3). Despite being able
to inhibit the colonization of pathogen's spores on the leaves, the poplars
couldn't inhibit the growth of pathogen's spores at the end. While it is good
to inhibit the number of spore colonization, it is more important to inhibit
the growth of spore so that the spore do not cover the leaves. This is because
if the poplars allow the spore to cover the leaf area, the poplars are at risk
to reduced growth due to inefficient photosynthesize.

---

# Obj 3 Observing the origins of poplar individuals collected from parent population and natural population
```{r bar chart classe, fig.height=5, fig.width=7}
x_labs = c("Natural Population", "Parent Population")

ggplot(wild_yr2000_new, aes(x = Classe)) +
  geom_bar(aes(fill = Classe)) + 
  geom_text(stat = "count", aes(label = ..count..), vjust = -0.5) +
  scale_x_discrete(labels = x_labs) +
  ggpubr::fill_palette("jco") +
  labs(x = NULL, y = "Counts", title = "The types of black poplar individuals") +
  theme_bw() +
  theme(panel.grid.major = element_blank(),
        legend.position = "none")
```

The poplar individuals were consisted of poplars collected from natural population and from parent population.  

Parent population means that these poplars have been used in a breeding program. Poplars are managed as hybrids. 
Hybrids are individuals created from 2 different species. In this case, hybrid poplar can be created from parents
Populus deltoides x Populus nigra (P. nigra is the scientific name of black poplar). So poplar individuals collected from parent population means that these individuals are used as a parent in breeding hybrid poplars.

So let's see their origins.

## Obj 3.1 The origins of poplar from natural and parent population
```{r origin natural vs parent}
# Insert a new column for "Region"
library(tibble)
wild_copy = wild_yr2000_new

wild_copy = tibble::add_column(wild_copy, "Region" = "NA", .after = "Population")

# Assigning the population to their region
for (j in c(1:nrow(wild_copy))) {
  if(wild_copy$Population[j] == "ValdeLoire" | 
     wild_copy$Population[j] == "Selles_sur_Cher" |
    wild_copy$Population[j] == "Bonny"){
    wild_copy$Region[j] = "Centre_Val_de_Loire"
  } 
  
  if(wild_copy$Population[j] == "Ramières" | 
     wild_copy$Population[j] == "ValAllier" |
     wild_copy$Population[j] == "St_Jean_de_Maurienne" |
     wild_copy$Population[j] == "Meximieux" | 
     wild_copy$Population[j] == "St_Genix_sur_Guiers"){
    wild_copy$Region[j] = "Auvergne_Rhone_Alpes"
  } 
  
  if(wild_copy$Population[j] == "Tarbes" | 
     wild_copy$Population[j] == "Nohèdes" |
     wild_copy$Population[j] == "Sauveterre" |
     wild_copy$Population[j] == "Cauterets"){
    wild_copy$Region[j] = "Occitanie"
  }
  
  if(wild_copy$Population[j] == "StPrivé" | 
     wild_copy$Population[j] == "La_Celle_sur_Loire"){
    wild_copy$Region[j] = "Bourgogne_Franche_Comte"
  }
  
  if(wild_copy$Population[j] == "Sarrazin" | 
     wild_copy$Population[j] == "Gave-Mauleon" |
     wild_copy$Population[j] == "Baudreix" |
     wild_copy$Population[j] == "Baliros" | 
     wild_copy$Population[j] == "Assat"){
    wild_copy$Region[j] = "Nouvelle_Aquitaine"
  }
  
  if(wild_copy$Population[j] == "Pont_de_Cé" | 
     wild_copy$Population[j] == "Cunault"){
    wild_copy$Region[j] = "Pays_de_la_Loire"
  }
  
  if(wild_copy$Population[j] == "Cazebonne"){
    wild_copy$Region[j] = "Bordeaux"
  }
  
}

lab_names = c("Natural Population", "Parent Population")
names(lab_names) = c("nigrapop", "parent 14x13")

x_labels = c("Auvergne Rhône Alpes", "Bordeaux",
             "Bourgogne Franche Comté", "Centre-Val de Loire",
             "Unknown", "Nouvelle Aquitaine",
             "Occitanie", "Pays de la Loire")

str(wild_copy)

wild_copy$Region = factor(wild_copy$Region)
  
ggplot(wild_copy, aes(x = Region)) +
  geom_bar(aes(fill = Classe)) +  # barchart
  ggpubr::fill_palette("jco") +
  coord_flip() +
  scale_x_discrete(labels = x_labels) +
  facet_wrap(~ Classe, labeller = labeller(Classe = lab_names), ncol = 1) +  # faceted by parents and natural pop
  labs( x = NULL, y = "Counts", title = "Population origins of black poplar individuals" ) +
  geom_text(stat = "count", aes(label = ..count..), hjust = -0.15, size = 3) +
  theme_bw() +
  theme(axis.text = element_text(size = 10),
        legend.position = "none")
```

## Obj 3.2 Comparison of resistance between poplars from natural population and from parent population
In this section, I would like to show you the resistance between the two populations. 
Are poplars from natural population resisting better to leaf rust than the poplars from parent population?

### Field resistance and spore size
```{r comparison resistance, fig.height=4, fig.width=8}
# Summary statistics with aggregate function

# Length
aggregate = aggregate(wild_yr2000_new$Classe, by = list(Classe = wild_yr2000_new$Classe), 
                       FUN = length)
names(aggregate)[2] = "length"

# Function mean
mean.fun = function(x){
  aggregate(x,
            by = list(Classe = wild_yr2000_new$Classe),
            FUN = mean)
}

# Function sd 
sd.fun = function(x){
  aggregate(x,
            by = list(Classe = wild_yr2000_new$Classe),
            FUN = sd)
}

# Plotting the comparison
compare_NM = ggplot(wild_yr2000_new, aes(x = Classe, y = Notemax.2010)) +
  geom_boxplot(aes(fill= Classe)) +
  ggpubr::fill_palette("jco") + 
  stat_summary(fun = "mean", geom = "point", size = 3, color = "red", shape = 15) +
  geom_jitter(color = "grey", width = .1) +
  labs(title = "Field resistance",
       x = "Class",
       y = "Score") +
  scale_x_discrete(limit = c("nigrapop", "parent 14x13"),
                   labels = c("Natural Population", "Parent Population")) +
  ylim(1, 6) +
  theme_classic() +
  theme(legend.position = "none")

# Spore size
compare_SS = ggplot(wild_yr2000_new, aes(x = Classe, y = taille.93JE3.2010)) +
  geom_boxplot(aes(fill = Classe)) +
  ggpubr::fill_palette("jco") +
  stat_summary(fun = "mean", geom = "point", size = 3, color = "red", shape = 15) +
  geom_jitter(color = "grey", width = .1) +
  labs(title = "Spore size",
       x = "Class",
       y = "Score") +
  scale_x_discrete(limit = c("nigrapop", "parent 14x13"),
                   labels = c("Natural Population", "Parent Population")) +
  ylim(1, 6) +
  theme_classic() +
  theme(legend.position = "none")

# Combining note max and spore size distribution
compare = ggarrange(compare_NM, compare_SS)
compare
```

In terms of field resistance:  
The poplars from natural populations show lower field resistance than parent population.  

In terms of spore size in the lab:  
The poplars from natural populations have lower spore size than the parents (on average).
This shows the ability of the poplars from natural populations to better prevent
the development of pathogen's spores on their leaves, **in laboratory setting**. 
It also means that in the lab, P. nigra itself is able to prevent the development 
of the spore, rather than when it is paired with other Populus species in a breeding
program.

### Latent period
```{r comparison LP, fig.height=5, fig.width=7}
# Latent period
compare_LP = ggplot(wild_yr2000_new, aes(x = Classe, y = latence.93JE3.2010)) +
  geom_boxplot(aes(fill = Classe)) +
  ggpubr::fill_palette("jco") +
  stat_summary(fun = "mean", geom = "point", size = 3, color = "red", shape = 15) +
  geom_jitter(color = "grey", width = .1) +
  labs(title = "Latent period",
       x = "Class",
       y = "Days") +
  scale_x_discrete(limit = c("nigrapop", "parent 14x13"),
                   labels = c("Natural Population", "Parent Population")) +
  ylim(7, 12) +
  theme_classic() +
  theme(legend.position = "none")

compare_LP
```

In terms of latent period in the lab:
The poplars from natural populations start showing symptoms in a much later day
than the poplars from the parent population. It shows the ability of the poplars from natural populations to 
prevent the accumulation of pathogen on the leaves.

When the poplars are able to prevent the accumulation of the pathogen, they are more
likely to prevent the growth of the pathogen on the leaves. Thus when they start showing
symptoms in a much later day, they tend to have fewer spores with smaller size on
the leaves than the other poplars that show symptoms in an earlier day.

### Spore number
```{r comparison NS, fig.height=5, fig.width=7}
# Number of spores
compare_NS = ggplot(wild_yr2000_new, aes(x = Classe, y = nbsores.93JE3.2010)) +
  geom_boxplot(aes(fill = Classe)) +
  ggpubr::fill_palette("jco") +
  stat_summary(fun = "mean", geom = "point", size = 3, color = "red", shape = 15) +
  geom_jitter(color = "grey", width = .1) +
  labs(title = "Number of spores",
       x = "Class",
       y = "Number (counts)") +
  scale_x_discrete(limit = c("nigrapop", "parent 14x13"),
                   labels = c("Natural Population", "Parent Population")) +
  theme_classic() +
  theme(legend.position = "none")

compare_NS
```

In terms of number of spores:
The poplars from natural populations have fewer number of spores than the parent.
It relates to the inhibiting actions of latent period, and it also leads to the smaller
spore size.

---

### Bottom line
So the bottom line is, poplars from natural population show lower resistance in the field than
the poplars from parent population. However, in the laboratory setting (a controlled environment),
the poplars from natural population show better resistance than those from parent population, in a way
that they are able to inhibit the growth of the pathogen.

---

# Obj 3 Correlation between field resistance and lab resistance
To observe the relationships between the parameters of resistance. Furthermore, 
to see if field resistance can be explained by the parameters of lab resistance
or not.
```{r correlation, message=FALSE, warning=FALSE}


  
# remove (?) --> let's see later
# sort from the population with the highest count
# vertical text *checked
# split between parents and natural pop (facet) *checked
# add data label *checked


#(resist_field = ggplot(wild_yr2000_clean, aes(x = Res_vs_sus_note)) +
  # geom_bar(aes(fill = Res_vs_sus_note)) +
  # geom_text(stat = "count", aes(label = ..count..), vjust = -0.5) +
  # labs(title = "Resistant vs susceptible poplars",
       # x = "Resistant vs Susceptibe",
       # y = "Counts") +
  # scale_fill_manual(values = c("#87CEEB", "#FF6A6A"),
                    # breaks = c("Resistant", "Susceptible"),
                    # name = "Resistant vs Susceptible",
                    # labels = c("Resistant", "Susceptible")) +
  #theme_classic() +
  #theme(axis.title.x = element_blank()))
LP_NM = ggpubr::ggscatter(wild_yr2000_new, 
                          x = "latence.93JE3.2010", y = "Notemax.2010",
                          add = "reg.line", conf.int = TRUE,
                          cor.coef = TRUE, cor.method = "pearson",
                          xlab = "Latent Period", ylab = "Field Resistance")

NS_NM = ggpubr::ggscatter(wild_yr2000_new, 
                          x = "nbsores.93JE3.2010", y = "Notemax.2010",
                          add = "reg.line", conf.int = TRUE,
                          cor.coef = TRUE, cor.method = "pearson",
                          xlab = "Number of spores", ylab = "Field Resistance")

SS_NM = ggpubr::ggscatter(wild_yr2000_new, 
                          x = "taille.93JE3.2010", y = "Notemax.2010",
                          add = "reg.line", conf.int = TRUE,
                          cor.coef = TRUE, cor.method = "pearson",
                          xlab = "Spore Size", ylab = "Field Resistance")

corr = ggpubr::ggarrange(LP_NM, NS_NM, SS_NM)

corr
```

It seems that there are no clear correlation between field and lab resistance. 
It seems unclear whether lab resistance can explain or predict field resistance.

# Obj 4 Correlation between parameters of lab resistance

```{r corr LP x SS, message=FALSE, warning=FALSE}
# Designing the main correlation plot latence x taille
(main_cor = ggpubr::ggscatter(wild_yr2000_new,
                              x = "latence.93JE3.2010", y = "taille.93JE3.2010",
                              add = "reg.line", conf.int = TRUE,
                              color = "Classe", palette = "jco",
                              shape = "Classe", size = 2,
                              alpha = 0.6, ggtheme = theme_bw(),
                              xlab = "Latent period (days)", 
                              ylab = "Spore size (score)") +
  stat_cor(aes(color = Classe)))  # correlation line

# Because I differentiated correlation for parent and nigrapop, I want to provide
# the density histogram for poplars from parent and nigrapop

# Density histogram for latent period, plotted at x axis
xdens = axis_canvas(main_cor, axis = "x") +  # create the canvas base
  geom_density(data = wild_yr2000_new, 
               aes(x = latence.93JE3.2010, fill = Classe),
               alpha = 0.7, size = 0.2) +  # density histogram for latence
  ggpubr::fill_palette("jco")  # color palette for fill 

# Density histogram for taille, plotted at y axis
ydens = axis_canvas(main_cor, axis = "y", coord_flip = TRUE) +
  geom_density(data = wild_yr2000_new, 
               aes(x = taille.93JE3.2010, fill = Classe),
               alpha = 0.7, size = 0.2) +  # density histogram for taille
  coord_flip() +  # flipping the coordinate, 90 degrees
  ggpubr::fill_palette("jco")


# Adding xdens on the x axis of correlation plot
main_cor1 = cowplot::insert_xaxis_grob(main_cor, xdens,
                                       grid::unit(0.2, "null"), 
                                       position = "top")

# Adding ydens on the y axis of correlation plot
main_cor2 = cowplot::insert_yaxis_grob(main_cor1, ydens,
                                       grid::unit(0.2, "null"), 
                                       position = "right")

cowplot::ggdraw(main_cor2)

# Saving the plot
ggsave("correlation_plot.png",
       device = "png",
       width = 15,
       height = 10,
       units = "cm",
       dpi = 300)
```

It seems that there is a clear correlation between the parameters of lab resistance
(latent period and spore size).