---
title: "Data cleaning"
---

```{r libraries, message=FALSE, warning=FALSE, include=FALSE}
library(readxl)  # to import the data in xlsx file
library(finalfit)  # for function ffglimpse(), extension of summary
library(ggplot2)  # for plotting
library(plyr)  # for counting the frequency from histogram
library(ggpubr)  # to get the function ggarrange; to combine ggplots
library(cowplot)  # to draw density histogram at the sides of correlation plot

# The dataset
pop_raw = readxl::read_xlsx("Tested_Poplar.xlsx",
                            na = "NA")  # defining the code for missing data used in the file
# Getting only the important variables
wild_yr2000 = pop_raw[,c("Classe",  # the type of population source 
                         "Population",  # population of origin
                         "Clone",  # the observed poplar individuals
                         "Notemax.2010 Noveltree",  # field resistance 
                         "latence.93JE3.2010",   # latent period
                         "nbsores.93JE3.2010",  # number of spores
                         "taille.93JE3.2010")]  # size of spores

colnames(wild_yr2000)[4] = "Notemax.2010"  # because the name was too long

# Check the class of each variable
str(wild_yr2000)

# Coercing
## Classe = chr to factor
## Population = chr to factor
## Clone = chr to factor
cols = c("Classe", "Population", "Clone")  # cols to be coerced

wild_yr2000[cols] = lapply(wild_yr2000[cols], factor)  # coerced to factor

# New data structure
str(wild_yr2000)
```

# Data cleaning: identifying NAs
```{r missing data}
# Examine NAs with ff_glimpse (package = finalfit)
explanatory = c("latence.93JE3.2010",
                "nbsores.93JE3.2010",
                "taille.93JE3.2010")

dependent = "Notemax.2010"

wild_yr2000 %>%
  finalfit::ff_glimpse(dependent, explanatory)
```

1. Missing percent for Notemax = 12.9 (ratio = 12/93)  
2. Missing percent for latence = 7.5 (ratio = 7/93)  
3. Missing percent for nbsores = 0.0 (ratio = 0/93)  
4. Missing percent for taille = 8.6 (ratio = 8/93)  

```{r plot missing data}
# Plotting the missing data
wild_yr2000 %>% 
  finalfit::missing_plot()

# NAs in notemax, latence, and taille
# When when latence is missing, taille is most likely to be missing as well
# Check if the missing data in taille is due to missing data in latence?

# Checking the relatedness of the missing data in these two variables 
# using missing_pairs
# testing H0 = missing points are MCAR (missing compeletely at random)
explanatory_1 = "latence.93JE3.2010"

dependent_1 = "taille.93JE3.2010"

wild_yr2000 %>% 
  finalfit::missing_compare(dependent_1, explanatory_1) %>%
  knitr::kable(row.names = FALSE, align = c("l", "l", "r", "r", "r"))

# P significant, missing data in taille is due to missing data in latence
```

The model (missing_pairs) shows that there is a significant difference in the mean of
spore size (taille) when there are missing data in latent period (latence) 
and when there are not.

> It means that the missing data in latence is related to the missing data in taille.

> Biologically speaking, it's how it usually is in the interaction between pathogen and 
the host. When latence is missing, it means that the leaves do not show symptoms during their
14 days in growth chamber (latence is non-measurable). When latence is non-measurable,
it is very likely that the number of spores and spore size are non-measurable as well
because the leaves do not show symptoms.

> This is also an indication that the leaves are resistant to the pathogen.

And this also means that there are systemic relationship among the lab resistance.

Solutions:  
1. To build a dashboard, we cannot exclude these missing data as they show 
the potential resistant poplars  
2. But we should exclude them when we model the relationship between field and lab
resistance  

---

What about between latence and field resistance?

```{r latence vs field missing}
explanatory_2 = "Notemax.2010"

dependent_2 = "latence.93JE3.2010"

wild_yr2000 %>% 
  finalfit::missing_compare(dependent_2, explanatory_2) %>%
  knitr::kable(row.names = FALSE, align = c("l", "l", "r", "r", "r"))
```

It seems that the missing data in latence do not relate to the missing data
in field resistance (MACR)

The missing data in notemax are also due to the missing infection on the poplar
leaves in the field. This shows that the leaves might not be infected by the pathogen,
showing the potential of the related poplar to be resistant. 

Therefore, these missing data should be included to build a dashboard for the share between
resistant and susceptible poplars.

# Data cleaning: Excluding controls (témoin) and poplars from national collection
To inspect resistant vs susceptible poplars, we have to exclude the 'discriminants'
or the controls (temoin). 

This is because they are just the controls for the infection in other poplars. We
have to clean the dataset from these controls to get a better look only on the poplars
from natural population and breeding population.

We should also exclude the poplars from national collection (collection nationale)
because it seems that these poplars were not tested in the field. Including these in 
linear regression will only create bias.
```{r exclude controls}
rownames_remove <- c("témoin", "coll. Nationale")

wild_yr2000_clean = subset(wild_yr2000,!(Classe %in% rownames_remove))
```