---
title: "Research Journal"
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
output: 
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 1
---

# Description

**About the raw dataset and the species used**

This analysis is to find the relationship between resistance measured in the field (field resistance) and measured in the lab (lab resistance).

The raw data consists of the measurement in field from year 2000 to 2010, as well as measurement in the lab for the same period.

The spesies is Populus nigra (poplar) that was collected from several natural populations and sources in Western Europe.

---

**Why poplar?**

Poplar is usually attacked by rust fungus that causes yellow spots on the poplar
leaves. These spots reduce poplar's effectiveness in carrying out photosynthesis processes
because these spots are covering the stomata in the leaves and reducing poplar's leaf
area to catch sunrays. The impact of the massive attack caused by rust fungus is a decreased
photosynthesis and growth rate which also lead to economic losses.

---

**Collecting resistance data**

The idea of collecting these resistance data is to find the resistant poplar that can
fight back the rust fungus.

*Field resistance*

Field resistance (Notemax) is measured in the field by scoring the infection on 
poplar leaves (score 1 - 6)

*Lab resistance*

Lab resistance is consisted of 3 parameters: latent period (latence), number of spores
on the leaves (nbsores), and spore size (taille). These parameters are measured in the
lab. The poplar leaves are firstly cut into disks and they are put in the leaf disks for
the lab experiment. The experiment then starts by infecting the poplar leaves with different
types of rust fungus (in the data set, they are coded like 98AG69, 93JE3, etc). 
Then they are put in the growth chamber for 14 days.

During 14 days, we observe:
1. Latent period = the time needed by the poplar leaves to start showing the first
symptom, measured daily  
2. Number of spores = number of spores found on the leaves, measured on the day
13 after infection  
3. Spore size = size of the spores, measured on the day 14 after inoculation using 
a scoring system (1 to 5, 1 being the smallest, 5 being the largest)  

---

# What's in the script
1. Raw dataset  
2. Identification of missing data (data cleaning)  
3. EDA and data dashboard  
4. Correlation plot  

---

# The objectives
1. Are there a completely resistant poplar in the field and in the lab?
2. What is the average resistance for poplar in the field and in the lab?
3. Is there a correlation between field resistance and lab resistance?
4. Is there a correlation between the parameters of lab resistance?

---

# About the raw dataset
Poplar species used in this dataset is Populus nigra
Raw dataset contains:  

* Classe = classes of poplar's population source  
    + témoin = control  
    + parent 14x13 = poplar generated from other breeding program  
    + nigrapop = poplar collected from natural populations  
    + coll. nationale = poplar collected from (French) national collection)  

* Population = the name of the population source  

* Department = the region where the population is located  
* Clone = the assigned name for each observed poplar  
* Stabilité intersouche and Nb plants en serre41 = additional information (not considered as variables)  
* Notemax = poplar infection measured in the field using scoring system (1 - 6), the year following "Notemax" indicates the year it was measured, and city names following the year indicates the location it was measured  
* Latence = latent period (the time needed by poplars to start showing the first symptoms, observed in days in a lab experiment)  
* Nbsores = number of spores on poplar leaves, measured in counts  
* Taille = size of spores on poplar leaves, measured using scoring system 1 - 5 (1 being the smallest, 5 being the largest) Codes following "Latence", "nbsores", and "taille" indicate the types of pathogen that infect poplar. The year following it indicates the year it was measured  

# Libraries
```{r libraries, message=FALSE, warning=FALSE}
library(readxl)  # to import the data in xlsx file
library(finalfit)  # for function ffglimpse(), extension of summary
library(ggplot2)  # for plotting
library(plyr)  # for counting the frequency from histogram
library(ggpubr)  # to get the function ggarrange; to combine ggplots
library(cowplot)  # to draw density histogram at the sides of correlation plot
```

# The dataset
```{r import raw}
pop_raw = readxl::read_xlsx("Tested_Poplar.xlsx",
                            na = "NA")  # defining the code for missing data used in the file
# Getting only the important variables
wild_yr2000 = pop_raw[,c("Classe",  # the type of population source 
                         "Population",  # population of origin
                         "Clone",  # the observed poplar individuals
                         "Notemax.2010 Noveltree",  # field resistance 
                         "latence.93JE3.2010",   # latent period
                         "nbsores.93JE3.2010",  # number of spores
                         "taille.93JE3.2010")]  # size of spores

colnames(wild_yr2000)[4] = "Notemax.2010"  # because the name was too long

# Check the class of each variable
str(wild_yr2000)

# Coercing
## Classe = chr to factor
## Population = chr to factor
## Clone = chr to factor
cols = c("Classe", "Population", "Clone")  # cols to be coerced

wild_yr2000[cols] = lapply(wild_yr2000[cols], factor)  # coerced to factor

# New data structure
str(wild_yr2000)
```

# Raw distribution
```{r raw distribution}
par(mfrow = c(2,2))  # combine the graphs (2 columns, 2 rows)

# Create a dataframe for loop
raw_loop = data.frame("note" = wild_yr2000$Notemax.2010,
                      "latence" = wild_yr2000$latence.93JE3.2010,
                      "nbsores" = wild_yr2000$nbsores.93JE3.2010,
                      "taille" = wild_yr2000$taille.93JE3.2010)

for (i in 1:4) {  # 1:4 for the number of variables used
  x = raw_loop[,i]  # assign x for each iteration
  title = names(raw_loop[i])  # title for each iteration
  hist(x,
       main = title)  # create histogram in each iteration
}
```

Seems that there are no outliers. 

Latent period is not normally distributed, which is fine, biologically speaking.
It shows that there is greater amount of poplars that start showing symptoms in early time
than in later time. Showing that most poplars are infected early. Meaning that
they are susceptible to the pathogen in the lab.