---
title: "Relationship between field and lab resistance"
author: "Firza Riany"
date: "12/11/2020"
output: 
  html_document:
    toc: TRUE
---

# Description
Read readme for details 

# Libraries
```{r libraries, message=FALSE, warning=FALSE}
library(readxl)  # to import the data in xlsx file
library(finalfit)  # for function ffglimpse(), extension of summary
library(ggplot2)  # for plotting
library(plyr)  # for counting the frequency from histogram
library(ggpubr)  # to get the function ggarrange; to combine ggplots
library(cowplot)  # to draw density histogram at the sides of correlation plot
```

# The dataset
```{r import raw}
pop_raw = readxl::read_xlsx("Tested_Poplar.xlsx",
                            na = "NA")  # defining the code for missing data used in the file
# Getting only the important variables
wild_yr2000 = pop_raw[,c("Classe",  # the type of population source 
                         "Population",  # population of origin
                         "Clone",  # the observed poplar individuals
                         "Notemax.2010 Noveltree",  # field resistance 
                         "latence.93JE3.2010",   # latent period
                         "nbsores.93JE3.2010",  # number of spores
                         "taille.93JE3.2010")]  # size of spores

colnames(wild_yr2000)[4] = "Notemax.2010"  # because the name was too long

# Check the class of each variable
str(wild_yr2000)

# Coercing
## Classe = chr to factor
## Population = chr to factor
## Clone = chr to factor
cols = c("Classe", "Population", "Clone")  # cols to be coerced

wild_yr2000[cols] = lapply(wild_yr2000[cols], factor)  # coerced to factor

# New data structure
str(wild_yr2000)
```

# Raw distribution
```{r raw distribution}
par(mfrow = c(2,2))  # combine the graphs (2 columns, 2 rows)

# Create a dataframe for loop
raw_loop = data.frame("note" = wild_yr2000$Notemax.2010,
                      "latence" = wild_yr2000$latence.93JE3.2010,
                      "nbsores" = wild_yr2000$nbsores.93JE3.2010,
                      "taille" = wild_yr2000$taille.93JE3.2010)

for (i in 1:4) {  # 1:4 for the number of variables used
  x = raw_loop[,i]  # assign x for each iteration
  title = names(raw_loop[i])  # title for each iteration
  hist(x,
       main = title)  # create histogram in each iteration
}
```

Seems that there are no outliers. 

Latent period is not normally distributed, which is fine, biologically speaking.
It shows that there is greater amount of poplars that start showing symptoms in early time
than in later time. Showing that most poplars are infected early. Meaning that
they are susceptible to the pathogen in the lab.

# Data cleaning: identifying NAs
```{r missing data}
# Examine NAs with ff_glimpse (package = finalfit)
explanatory = c("latence.93JE3.2010",
                "nbsores.93JE3.2010",
                "taille.93JE3.2010")

dependent = "Notemax.2010"

wild_yr2000 %>%
  finalfit::ff_glimpse(dependent, explanatory)
```

1. Missing percent for Notemax = 12.9 (ratio = 12/93)  
2. Missing percent for latence = 7.5 (ratio = 7/93)  
3. Missing percent for nbsores = 0.0 (ratio = 0/93)  
4. Missing percent for taille = 8.6 (ratio = 8/93)  

```{r plot missing data}
# Plotting the missing data
wild_yr2000 %>% 
  finalfit::missing_plot()

# NAs in notemax, latence, and taille
# When when latence is missing, taille is most likely to be missing as well
# Check if the missing data in taille is due to missing data in latence?

# Checking the relatedness of the missing data in these two variables 
# using missing_pairs
# testing H0 = missing points are MCAR (missing compeletely at random)
explanatory_1 = "latence.93JE3.2010"

dependent_1 = "taille.93JE3.2010"

wild_yr2000 %>% 
  finalfit::missing_compare(dependent_1, explanatory_1) %>%
  knitr::kable(row.names = FALSE, align = c("l", "l", "r", "r", "r"))

# P significant, missing data in taille is due to missing data in latence
```

The model (missing_pairs) shows that there is a significant difference in the mean of
spore size (taille) when there are missing data in latent period (latence) 
and when there are not.

It means that the missing data in latence is related to the missing data in taille.

Biologically speaking, it's how it usually is in the interaction between pathogen and 
the host. When latence is missing, it means that the leaves do not show symptoms during their
14 days in growth chamber (latence is non-measurable). When latence is non-measurable,
it is very likely that the number of spores and spore size are non-measurable as well
because the leaves do not show symptoms.

This is also an indication that the leaves are resistant to the pathogen.

And this also means that there are systemic relationship among the lab resistance.

Solutions:
1. To build a dashboard, we cannot exclude these missing data as they show 
the potential resistant poplars
2. But we should exclude them when we model the relationship between field and lab
resistance  

What about between latence and field resistance?

```{r latence vs field missing}
explanatory_2 = "Notemax.2010"

dependent_2 = "latence.93JE3.2010"

wild_yr2000 %>% 
  finalfit::missing_compare(dependent_2, explanatory_2) %>%
  knitr::kable(row.names = FALSE, align = c("l", "l", "r", "r", "r"))
```

It seems that the missing data in latence do not relate to the missing data
in field resistance (MACR)

The missing data in notemax are also due to the missing infection on the poplar
leaves in the field. This shows that the leaves might not be infected by the pathogen,
showing the potential of the related poplar to be resistant. 

Therefore, these missing data should be included to build a dashboard for the share between
resistant and susceptible poplars.

# Data cleaning: Excluding controls (témoin) and poplars from national collection
To inspect resistant vs susceptible poplars, we have to exclude the 'discriminants'
or the controls (temoin). 

This is because they are just the controls for the infection in other poplars. We
have to clean the dataset from these controls to get a better look only on the poplars
from natural population and breeding population.

We should also exclude the poplars from national collection (collection nationale)
because it seems that these poplars were not tested in the field. Including these in 
linear regression will only create bias.
```{r exclude controls}
rownames_remove <- c("témoin", "coll. Nationale")

wild_yr2000_clean = subset(wild_yr2000,!(Classe %in% rownames_remove))
```

# Obj 1 Data dashboard for resistant vs susceptible (in the field)
Turned out the NA is due to
1. Non-measured field resistance in the poplars from national collection.  
2. Non-active infection in certain controls.  

```{r res vs sus field}
# Resistant poplars in the field: Assigning -5 to NA to avoid confusion in R
wild_yr2000_clean[is.na(wild_yr2000_clean)] = -5

# Add a new variable: resistant vs susceptible
# Loop, -5 = resistant, != -5 susceptible
wild_yr2000_clean[,8] = "NA"

names(wild_yr2000_clean)[8] = "Res_vs_sus_note"

for (i in c(1:nrow(wild_yr2000_clean))) {
  if(wild_yr2000_clean$Notemax.2010[i] == -5)
    wild_yr2000_clean[i,8] = "Resistant"
  
  if(wild_yr2000_clean$Notemax.2010[i] != -5)
    wild_yr2000_clean[i,8] = "Susceptible"
}

# Plotting resistant vs susceptible poplars in terms of field resistance
(resist_field = ggplot(wild_yr2000_clean, aes(x = Res_vs_sus_note)) +
  geom_bar(aes(fill = Res_vs_sus_note)) +
  geom_text(stat = "count", aes(label = ..count..), vjust = -0.5) +
  labs(title = "Resistant vs susceptible poplars",
       x = "Resistant vs Susceptibe",
       y = "Counts") +
  scale_fill_manual(values = c("#87CEEB", "#FF6A6A"),
                    breaks = c("Resistant", "Susceptible"),
                    name = "Resistant vs Susceptible",
                    labels = c("Resistant", "Susceptible")) +
  theme_classic() +
  theme(axis.title.x = element_blank()))

# Checking the population of origin of the resistant poplars
pop = plyr::count(wild_yr2000_clean$Population[
  which(wild_yr2000_clean$Notemax.2010 == -5)
])

names(pop)[1] = "Population"

clone = plyr::count(wild_yr2000_clean$Clone[
  which(wild_yr2000_clean$Notemax.2010 == -5)
])

names(clone)[1] = "Clone"

pop$Clone = paste(clone$Clone)

pop

```

It seems that there is only one poplar (71072-501) from St Genix that is resistant 
in the field.

Let's see the resistant poplars in terms of lab resistance
Those with latence = 14 and taille = -5

```{r res vs sus lab}
# Add a new variable: resistant vs susceptible (lab)
# Loop, -5 = resistant, != -5 susceptible
wild_yr2000_clean[,9] = "NA"

names(wild_yr2000_clean)[9] = "Res_vs_sus_lab"

for (i in c(1:nrow(wild_yr2000_clean))) {
  if(wild_yr2000_clean$taille.93JE3.2010[i] == -5
     & wild_yr2000_clean$latence.93JE3.2010[i] == 14)
    wild_yr2000_clean[i,9] = "Resistant"
  
  if(wild_yr2000_clean$taille.93JE3.2010[i] != -5
     & wild_yr2000_clean$latence.93JE3.2010[i] != 14)
    wild_yr2000_clean[i,9] = "Susceptible"
}

# Plotting resistant vs susceptible poplars in terms of lab resistance
(resist_lab = ggplot(wild_yr2000_clean, aes(x = Res_vs_sus_lab)) +
  geom_bar(aes(fill = Res_vs_sus_lab)) +
  geom_text(stat = "count", aes(label = ..count..), vjust = -0.5) +
  labs(title = "Resistant vs susceptible poplars",
       x = "Resistant vs Susceptibe",
       y = "Counts") +
  scale_fill_manual(values = c("#458B74", "#EE3B3B"),
                    breaks = c("Resistant", "Susceptible"),
                    name = "Resistant vs Susceptible",
                    labels = c("Resistant", "Susceptible")) +
  theme_classic() +
  theme(axis.title.x = element_blank()))

# Checking the population of origin of the resistant poplars (lab)
pop2 = plyr::count(wild_yr2000_clean$Population[
  which(wild_yr2000_clean$Res_vs_sus_lab == "Resistant")
])

names(pop2)[1] = "Population"

clone2 = plyr::count(wild_yr2000_clean$Clone[
  which(wild_yr2000_clean$Res_vs_sus_lab == "Resistant")
])

names(clone2)[1] = "Clone"

pop2$Clone = paste(clone2$Clone)

pop2
```

Seems that there are different poplars that are resistant in field and in the lab.

# Obj 2 Average resistance. Data dashboard for summary statistics
```{r summary statistics, fig.height=5, fig.width=10, message=FALSE, warning=FALSE}
# Field resistance (notemax)
# Removing -5 because they're outliers
wild_yr2000_new = wild_yr2000_clean[which(wild_yr2000_clean$Notemax.2010 != -5 &
                                      wild_yr2000_clean$taille.93JE3.2010 != -5), ]

summary(wild_yr2000_new)

NM = ggplot(wild_yr2000_new, aes(x = Notemax.2010)) +
  geom_histogram(binwidth = 0.5) +
  geom_density(aes(y = 0.3 * ..count..)) +
  geom_vline(aes(xintercept = mean(Notemax.2010), color = "Mean"), size = 1) +
  geom_vline(aes(xintercept = median(Notemax.2010), color = "Median"), size = 1) +
  geom_vline(aes(xintercept = quantile(Notemax.2010, prob = 0.25), color = "Q1"), 
             size = 1) +
  geom_vline(aes(xintercept = quantile(Notemax.2010, prob = 0.75), color = "Q3"), 
             size = 1) +
  labs(title = "Field resistance",
       caption = "number of samples = 72, sd = 0.63",
       x = "Score",
       y = "Counts") +
  scale_color_manual(name = "Statistics", values = c(Mean = "red", Median = "orange", Q1 = "blue",
                                                     Q3 = "blue")) +
  xlim(1, 6) +
  ylim(0, 40)
 
# Latent period (latence)
LP = ggplot(wild_yr2000_new, aes(x = latence.93JE3.2010)) +
  geom_histogram(binwidth = 0.5) +
  geom_density(aes(y = 0.3 * ..count..)) +
  geom_vline(aes(xintercept = mean(latence.93JE3.2010), color = "Mean"), size = 1) +
  geom_vline(aes(xintercept = median(latence.93JE3.2010), color = "Median"), size = 1) +
  geom_vline(aes(xintercept = quantile(latence.93JE3.2010, prob = 0.25), color = "Q1"), 
             size = 1) +
  geom_vline(aes(xintercept = quantile(latence.93JE3.2010, prob = 0.75), color = "Q3"), 
             size = 1) +
  labs(title = "Latent period",
       caption = "sd = 0.69",
       x = "Days",
       y = "Counts") +
  scale_color_manual(name = "Statistics", values = c(Mean = "red", Median = "orange", Q1 = "blue",
                                                     Q3 = "blue")) +
  xlim(7, 12) +
  ylim(0, 40)

# sd(wild_yr2000_new$latence.93JE3.2010)

# Number of spores
NS = ggplot(wild_yr2000_new, aes(x = nbsores.93JE3.2010)) +
  geom_histogram(binwidth = 1.0) +
  geom_density(aes(y = 0.3 * ..count..)) +
  geom_vline(aes(xintercept = mean(nbsores.93JE3.2010), color = "Mean"), size = 1) +
  geom_vline(aes(xintercept = median(nbsores.93JE3.2010), color = "Median"), size = 1) +
  geom_vline(aes(xintercept = quantile(nbsores.93JE3.2010, prob = 0.25), color = "Q1"), 
             size = 1) +
  geom_vline(aes(xintercept = quantile(nbsores.93JE3.2010, prob = 0.75), color = "Q3"), 
             size = 1) +
  labs(title = "Number of spores",
       caption = "sd = 9.49",
       x = "Number (counts)",
       y = "Counts") +
  scale_color_manual(name = "Statistics", values = c(Mean = "red", Median = "orange", Q1 = "blue",
                                                     Q3 = "blue")) +
  xlim(0, 60) +
  ylim(0, 40)

# sd(wild_yr2000_new$nbsores.93JE3.2010)

# Spore size
SS = ggplot(wild_yr2000_new, aes(x = taille.93JE3.2010)) +
  geom_histogram(binwidth = 0.5) +
  geom_density(aes(y = 0.3 * ..count..)) +
  geom_vline(aes(xintercept = mean(taille.93JE3.2010), color = "Mean"), size = 1) +
  geom_vline(aes(xintercept = median(taille.93JE3.2010), color = "Median"), size = 1) +
  geom_vline(aes(xintercept = quantile(taille.93JE3.2010, prob = 0.25), color = "Q1"), 
             size = 1) +
  geom_vline(aes(xintercept = quantile(taille.93JE3.2010, prob = 0.75), color = "Q3"), 
             size = 1) +
  labs(title = "Spore size",
       caption = "sd = 0.68",
       x = "Score",
       y = "Counts") +
  scale_color_manual(name = "Statistics", values = c(Mean = "red", Median = "orange", Q1 = "blue",
                                                     Q3 = "blue")) +
  xlim(1, 5) +
  ylim(0, 40)

# sd(wild_yr2000_new$taille.93JE3.2010)
quantile(wild_yr2000_new$nbsores.93JE3.2010, prob = 0.75)

# Combining the plots
dist = ggpubr::ggarrange(NM, LP, NS, SS)

dist
```

* Field resistance  
Seems that the poplars are in intermediate resistance, on average. For a note,
field resistance is measured using scoring system from 1 - 6; score 1 means
that infection covers only a small area of leaves, score 6 means that infection
covers almost all area of poplar leaves. In other words, when field resistance
score is 1 then the poplar is resistant, when field resistance score is 6 then
the poplar is susceptible.  
  
  In the field, the score is a little bit more than 3, which means that the infection
moderately covers poplar's leaf area. This being said that the poplars are either
resistant or susceptible in the field.

* Lab resistance, parameter: latent period  
On average, poplars started showing the first infection (or symptoms) on day
8.5 after inoculation. Day 8.5 means that the infection starts on day 8 after 
midday. Seems that the poplars can inhibit the colonization of the pathogen 
until day 8, on average.

* Lab resistance, parameter: number of spores  
Poplar individuals show a variable number of spores. Because of this, there are 
not many individuals that have similar number of spores (see the Y-axis scale). 
It seems that there are 2 peaks in the distribution, showing the segregation
between indivdiuals. This is the indication that the population is made of 2 groups
of populations. On average, the poplars are colonized by 30 spores (max = 56 spores)
even though there are not so many poplars colonized by 30 spores. There are more
poplars colonized by either 25 or 36 spores, indicating that the poplars may be
moderately resistant.

* Lab resistance, parameter: size of spores  
On average, the size of spores is scored closer to 4 than to 3. It shows that the
poplars cannot inhibit the development of the pathogen's spore once the spores
colonize the poplars. It shows that the poplars are rather susceptible for this 
type of pathogen (code: rust 93JE3)

* Overall  
The poplars are closer to being susceptible than to being resistant in the lab 
for this particular type of pathogen (code: rust 93JE3). Despite being able
to inhibit the colonization of pathogen's spores on the leaves, the poplars
couldn't inhibit the growth of pathogen's spores at the end. While it is good
to inhibit the number of spore colonization, it is more important to inhibit
the growth of spore so that the spore do not cover the leaves. This is because
if the poplars allow the spore to cover the leaf area, the poplars are at risk
to reduced growth due to inefficient photosynthesize.

# Obj 3 Correlation between field resistance and lab resistance
To observe the relationships between the parameters of resistance. Furthermore, 
to see if field resistance can be explained by the parameters of lab resistance
or not.

```{r correlation}
LP_NM = ggpubr::ggscatter(wild_yr2000_new, 
                          x = "latence.93JE3.2010", y = "Notemax.2010",
                          add = "reg.line", conf.int = TRUE,
                          cor.coef = TRUE, cor.method = "pearson",
                          xlab = "Latent Period", ylab = "Field Resistance")

NS_NM = ggpubr::ggscatter(wild_yr2000_new, 
                          x = "nbsores.93JE3.2010", y = "Notemax.2010",
                          add = "reg.line", conf.int = TRUE,
                          cor.coef = TRUE, cor.method = "pearson",
                          xlab = "Number of spores", ylab = "Field Resistance")

SS_NM = ggpubr::ggscatter(wild_yr2000_new, 
                          x = "taille.93JE3.2010", y = "Notemax.2010",
                          add = "reg.line", conf.int = TRUE,
                          cor.coef = TRUE, cor.method = "pearson",
                          xlab = "Spore Size", ylab = "Field Resistance")

corr = ggpubr::ggarrange(LP_NM, NS_NM, SS_NM)

corr
```

It seems that there are no clear correlation between field and lab resistance. 
It seems unclear whether lab resistance can explain or predict field resistance.

# Obj 4 Correlation between parameters of lab resistance

```{r corr LP x SS}
# Designing the main correlation plot latence x taille
(main_cor = ggpubr::ggscatter(wild_yr2000_new,
                              x = "latence.93JE3.2010", y = "taille.93JE3.2010",
                              add = "reg.line", conf.int = TRUE,
                              color = "Classe", palette = "jco",
                              shape = "Classe", size = 2,
                              alpha = 0.6, ggtheme = theme_bw(),
                              xlab = "Latent period (days)", 
                              ylab = "Spore size (score)") +
  stat_cor(aes(color = Classe)))  # correlation line

# Because I differentiated correlation for parent and nigrapop, I want to provide
# the density histogram for poplars from parent and nigrapop

# Density histogram for latent period, plotted at x axis
xdens = axis_canvas(main_cor, axis = "x") +  # create the canvas base
  geom_density(data = wild_yr2000_new, 
               aes(x = latence.93JE3.2010, fill = Classe),
               alpha = 0.7, size = 0.2) +  # density histogram for latence
  ggpubr::fill_palette("jco")  # color palette for fill 

# Density histogram for taille, plotted at y axis
ydens = axis_canvas(main_cor, axis = "y", coord_flip = TRUE) +
  geom_density(data = wild_yr2000_new, 
               aes(x = taille.93JE3.2010, fill = Classe),
               alpha = 0.7, size = 0.2) +  # density histogram for taille
  coord_flip() +  # flipping the coordinate, 90 degrees
  ggpubr::fill_palette("jco")


# Adding xdens on the x axis of correlation plot
main_cor1 = cowplot::insert_xaxis_grob(main_cor, xdens,
                                       grid::unit(0.2, "null"), 
                                       position = "top")

# Adding ydens on the y axis of correlation plot
main_cor2 = cowplot::insert_yaxis_grob(main_cor1, ydens,
                                       grid::unit(0.2, "null"), 
                                       position = "right")

cowplot::ggdraw(main_cor2)

# Saving the plot
ggsave("correlation_plot.png",
       device = "png",
       width = 15,
       height = 10,
       units = "cm",
       dpi = 300)
```

It seems that there is a clear correlation between the parameters of lab resistance
(latent period and spore size)